### 2009-05-25 doc/tech_doc/infosys/infosys_technical.tex
Technical documentation corrections

\subsection{Functionality}
\label{sub:peer_to_peer_functionality}

\subsection{Interface}
\label{sub:isis_peer_to_peer_interface}

>>>>

\subsection{Functionality}
\label{sub:peer_to_peer_functionality}
The ISIS nodes are able to organize themselves into a peer-to-peer network by synchronizing the stored replicated database and maintain this synchronicity and the network topology. Every node of this peer-to-peer cloud (that is an ISIS service) has an own PeerID, an identifier only used in this network, made as a hash of its endpoint URL. The members are ordered by this identifier where there is also a sequence defined between them where the successor entity is the ISIS with the smallest PeerID of the greater PeerIDs. This metric is used for defining the neighbour relation that is the base of the inter-peer-to-peer network communication. It is presumed that only registered ISIS nodes can have any role in the cloud.

\subsubsection{Connecting to the peer-to-peer network}
Every ISIS tries to connect to the network at its startup by accessing the InfoProvider ISIS nodes known from its configuration. The choice from among the few pre-configured InfoProviders is random to achieve as much load-balancing as possible. (In the design it is assumed that a huge amount of ISIS nodes are configured in very similar way, where the InfoProvider ISIS nodes are mainly the same.) If there is no InfoProvider available or nothing is known from the configuration, the node decides to be the fist member and to build a new network.

If there are any InfoProvider nodes available the list of member ISIS nodes will be queried and the ÒsuccessorÓ will be defined using the previously mentioned metric. This ÒsuccessorÓ node will help the entering ISIS to connect to the cloud. The new member will use ÒConnectÓ Web Service operation to get the newest version of database stored in the peer-to-peer network, and place its own entries to the cloud by registering them. An initial database synchronization can be achieved using this method.

We can prevent the InfoProvider ISIS nodes from overloading by balancing the operation of database synchronizing being the most costly procedure in the connection phase.

There is also an other extension used regarding the InfoProvider ISIS nodes. If a node is not able to connect any of the InfoProviders then it tries to repeat it later. This results that the network is able to fuse two disjoint parts of network when the missing InfoProvider ISIS appears again.

\subsubsection{Routing in the peer-to-peer network}
We are keeping the databases in synch by passing every Registration and RemoveRegistrations to every node acting in the network. This yields redundancy and the fault tolerant message routing is very important between any two ISIS nodes.

The most important parameter of the network is the {\it sparsity}. This is an integer not less then 2. It determines the number of neighbors in the function of the actual number of member nodes of the network, where the number of neighbors can change when new members connecting to the network or members leave. A sparse graph can be shaped by choosing a great value for sparsity or a dense one by choosing the less. This value is defined during the design and determines the main properties of the peer-to-peer network.
\begin{figure}[ht]
\centering{{\scalebox{0.6}{\includegraphics{Circle2-1.png}}}
\caption{\label{fig:p2p_neighbors_2} The neighbor connections in the peer-to-peer network in case of sparsity=2} }
\end{figure}

Every node has always exactly n = $\lceil \log_s{N} \rceil$ neighbors where the $N$ is the number of ISIS nodes registered into the network. The node has as many neighbors as the ISIS nodes with PeerID $s^0$-th, $s^1$-th ... $s^{n-1}$-th greater than its own. (For example: There are 19 nodes in this network with sparsity=9. Then the first node has the following neighbors: the 2nd, 3rd, 5th, 9th, and 17th node as seen in figure \ref{fig:p2p_neighbors_2}.)

There can be exactly n-way redundancy achieved in the system so at most $n-1$ node can disappear without any serious communication impair. Since there is also an emergency method used trying to deliver messages even if all neighbors are unavailable, this fault tolerance is greater. With this extension, messages are passed to one of the non-neighbor, yet available ISIS, so the message will surely be delivered to one of the operational node of the network. This solution provides a secure and quite fast solution for the message delivery but has a high communication cost because there are $n*N$ messages necessary in general for every Registration or RemoveRegistrations.

If there is a much greater {\it sparsity} used, then the number of the expected members decrease, even only to one neighbor. This situation can be seen in Figure \ref{fig:p2p_neighbors_40}. In this case such functioning is achieved when there are only N (the theoretically minimum) messages traversing in the network at the expense of a slower data propagation.
\begin{figure}[ht]
\centering{{\scalebox{0.6}{\includegraphics{Circle40-40.png}}}
\caption{\label{fig:p2p_neighbors_40} The neighbor connections in the peer-to-peer network in case of sparsity>N} }
\end{figure}

There is a data-driven routing used in the peer-to-peer cloud. It means that the given nodes examine the received Registration Entries then store and forward only those messages that are newer than the already stored version belonging to the same service identifier. If the information in the message is out-of-date then it will be simple dropped. (A Coordinated Universal Time (UTC) standard time and date format is used for the time stamps that is also able to handle the differences originated from the different time zones.) By using this method it is achieved that it is not necessary to store the formerly seen nodes in the messages, but the routing decision that is who to send the message to is based on local information. Since the routing is based on time stamps, it is very important to keep the nodes clock in synch with each other or with an outer reference, say, by using NTP.

Another advantage of this solution is the capability of handling the case of swapped Register and RemoveRegistrations messages. It is possible for a RemoveRegistrations message to forerun a previously generated Register message. This causes just a temporary imprecise entry in the database but the mistake can be quickly fixed and is not much further propagated.

Since the applied routing is based on the state of the messages belonging to some service identifier it is important to keep the fact of message deleting for a while. If a RemoveRegistrations message is received, then the named entry from the database is not removed immediately, just its state changes to {\it deleted} and a piece of data about it (the service identifier and the deletion time) is kept.

\subsubsection{Maintenance of the peer-to-peer network}
In the network there are two different maintenance methods needed. On one hand the database have to be periodically cleaned and on the other hand the list of neighbors have always got to be up-to-date.

The database applied in the system is a soft-state database so the entries are {\it valid} just for a limited time. Without any registration freshening its validity will expire and the state of the entry will change to {\it expired}. The time of expiration is set in every singe message by the registrar entity. The RemoveRegistrations message turn the state of registration entry to {\it deleted} as it was previously mentioned. The {\it expired} or {\it deleted} messages will be purged in the course of periodical database maintenance. The states of registration entries are shown in Figure \ref{fig:RegEntry states}.
\begin{figure}[ht]
\centering{{\scalebox{0.6}{\includegraphics{RegEntryStates.png}}}
\caption{\label{fig:RegEntry states} Registration entry states} }
\end{figure}

The second kind of maintenance is keeping the list of neighbors up-to-date. While receiving even Registration or RemoveRegistrations messages or finding an outdated entry during the periodical database cleaning nodes perform necessary steps to update our list. Since every leaving or entering node infulences the neighbor registry of all the other nodes at a high probability the local set of neighbors will be rebuilt in these cases. The modifications are passing locally because these connections are asymmetric so there is no compliance or network communication is needed.

There is no special action done and no additional network traffic needed during the neighbor checking. The fact the node sends periodically messages to every neighbor can be exploited. If the node is not successful in sending its messages to the neighbors after some tries, then the unavailable node will be marked. If every neighbor has been marked then the node changes its PeerID and reconnects to the network if there are any InfoProvider ISIS nodes available.

When a neighbor is unavailable ISIS keeps trying to deliver the message to the successor of the neighbor until either the information is passed to at least one available node or it is diagnosed that all the nodes are unavailable. Here the successor means thhose nodes that have greater PeerID than the neighbor but less then the next ISIS neighbor.
% subsection functionality (end)

\subsection{Interface}
\label{sub:isis_peer_to_peer_interface}
\subsubsection{Operation Connect}

  \item[Output]~\begin{description}
    \item[RegEntry] Multiple \texttt{RegEntries} locally stored in the ISIS database
  \end{description}

This operation allows the ISIS to get the already existing database from the ISIS that helps it to connect the peer-to-peer network. As a second step of the connection the newer  \texttt{RegEntries} that are stored at the connecting entity will be propagated through \texttt{Register} operation of the standard service specific interface.
